---
title: "March Madness"
output:
  word_document: default
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

**FOR FUTURE REFERENCE**
**For any given year the only changes needed are the new 538 file, dat24 becomes dat~CURRENT_YEAR~ throughout, and regional for loop must be altered to represent the regions including the first four matchups. Everything else is seemless in theory.**


Download the probabilties data from Five Thirty Eight. (I manually formatted CSV to only be Mens Tournament on the back end)
Order data for the functions we create below. (Line 21)

```{r}
library(tidyverse)
library(mosaic)
library(dplyr)
library(ggplot2)
```


```{r}
dat <- read.csv("~/Desktop/madness24.csv", stringsAsFactors = FALSE)

dat24 <- dat
dat24[, 5:10] <- dat[, 5:10] / dat[, 4:9]

head(dat24)
view(dat24)
```
Create a general tournament function (the best team plays the worst team, the second best team plays the second worst team...)
So if you order the data frames as above,this makes picking the teams being simulated very easy (Line 34)
The line of code after that simulates a game by pulling one sample (or team) according to their win probability in a given round.
Collect our winners and output them as mat.small (Line 39)

```{r}
Tournament.Round <- function(mat.big = mat.16, rd = "rd2_win"){
  vec.small <- c()
  for(i in 1:(nrow(mat.big) / 2)){
    ind <- c(i, nrow(mat.big) + 1 - i)
    temp <- sample(x = ind, size = 1, prob = mat.big[ind, rd])
    vec.small <- c(vec.small, temp)
  }
  mat.small <- mat.big[vec.small, ]
}
```

Now we work by region in a for loop. This is nessecary to work through the BULLSHIT created by the First Four. 
For 2021 the If statments are only functioning on the West and East regions but for future reference you can see how it would be structured for a region with 0, 1 or 2 first four matchups. 

Then take the field of 64 and run through each following round based on the original probabilites. (Lines 94-103)

```{r}
Region.Simulation <- function(region = "Midwest", dat = dat24){
  mat.17 <- dat[which(dat24$team_region == region),]
  
  
  if(region == "West"){
    first.four <- which(mat.17$team_seed == "16a" | mat.17$team_seed == "16b")
    mat.16 <- rbind(mat.17[-first.four, ],
                    mat.17[sample(x = first.four, size = 1,
                                  prob = mat.17[first.four, "rd1_win"]), ])
    mat.16[16, "team_seed"] <- "16"
    mat.16[, "team_seed"] <- as.numeric(mat.16[, "team_seed"])
    mat.16 <- mat.16[order(mat.16$team_seed), ]
  }
  if(region == "East"){
    mat.16 <- mat.17
    mat.16[, "team_seed"] <- as.numeric(mat.16[, "team_seed"])
    mat.16 <- mat.16[order(mat.16$team_seed), ]
  }

  if(region == "Midwest"){
    first.four <- which(mat.17$team_seed == "16a" | mat.17$team_seed == "16b")
    mat.17 <- rbind(mat.17[-first.four, ],
                    mat.17[sample(x = first.four, size = 1,
                                  prob = mat.17[first.four, "rd1_win"]), ])
    mat.17[17, "team_seed"] <- "16"
    first.four <- which(mat.17$team_seed == "10a" | mat.17$team_seed == "10b")
    mat.16 <- rbind(mat.17[-first.four, ],
                    mat.17[sample(x = first.four, size = 1,
                                  prob = mat.17[first.four, "rd1_win"]), ])
    mat.16[16, "team_seed"] <- "10"

    mat.16[, "team_seed"] <- as.numeric(mat.16[, "team_seed"])
    mat.16 <- mat.16[order(mat.16$team_seed), ]
  }
  if(region == "South"){
    first.four <- which(mat.17$team_seed == "10a" | mat.17$team_seed == "10b")
    mat.16 <- rbind(mat.17[-first.four, ],
                    mat.17[sample(x = first.four, size = 1,
                                  prob = mat.17[first.four, "rd1_win"]), ])
    mat.16[16, "team_seed"] <- "10"
    mat.16[, "team_seed"] <- as.numeric(mat.16[, "team_seed"])
    mat.16 <- mat.16[order(mat.16$team_seed), ]
  }
  # Round of 16
  mat.8 <- Tournament.Round(mat.16, rd = "rd2_win")
  mat.4 <- Tournament.Round(mat.8, rd = "rd3_win")
  mat.2 <- Tournament.Round(mat.4, rd = "rd4_win")
  mat.1 <- Tournament.Round(mat.2, rd = "rd5_win")
  list(First_Four = mat.16[c(10, 16), "team_name"],
       First_Round = mat.8[, "team_name"],
       Second_Round = mat.4[, "team_name"],
       Regional_Semis = mat.2[, "team_name"],
       Regional_Final = mat.1[, c("team_name", "rd6_win", "rd7_win")])
}
```



Take the previous functions and use outputs to simulate the Final Four and the championship.

```{r}
March.Madness.Simulation <- function(dat = dat24){
  midwest <- Region.Simulation(region = "Midwest", dat = dat24)
  east <- Region.Simulation(region = "East", dat = dat24)
  west <- Region.Simulation(region = "West", dat = dat24)
  south <- Region.Simulation(region = "South", dat = dat24)
  reg1 <- sample(x = c("east", "west"), size = 1,
                 prob = c(east$Regional_Final$rd6_win,
                          west$Regional_Final$rd6_win))
  National_Semi1 <- get(reg1)$Regional_Final
  reg2 <- sample(x = c("midwest", "south"), size = 1,
                 prob = c(midwest$Regional_Final$rd6_win,
                          south$Regional_Final$rd6_win))
  National_Semi2 <- get(reg2)$Regional_Final
  National_Champ <- sample(x = c(National_Semi1$team_name,
                                 National_Semi2$team_name),
                           size = 1,
                           prob = c(National_Semi1$rd7_win,
                                    National_Semi2$rd7_win))
  midwest$Regional_Final <- midwest$Regional_Final$team_name
  east$Regional_Final <- east$Regional_Final$team_name
  west$Regional_Final <- west$Regional_Final$team_name
  south$Regional_Final <- south$Regional_Final$team_name
  
  list(Midwest = midwest, East = east, West = west, South = south,
       National_Semi = c(National_Semi1$team_name, National_Semi2$team_name),
       National_Champ = National_Champ)
}
 
```

Now the fun!! 
Create an empty vector and empty matrix of any replication size you want.(Lines 144-146)
For Loop runs the simulation with 1:ANY AMOUNT YOU DESIRE of replicates. (Lines 149-152)
From matrix to data frame. (Line 153)

```{r}
nat.champ.vec <- unlist(March.Madness.Simulation(dat = dat24))
nat.champ.mat <- matrix("", nrow = 10000, ncol = length(nat.champ.vec))
colnames(nat.champ.mat) <- names(nat.champ.vec)

for(i in 1:10000){
  set.seed(i)
  nat.champ.mat[i, ] <- unlist(March.Madness.Simulation(dat = dat24))
}
nat.champ.mat <- as.data.frame(nat.champ.mat) 
```

Sort the data frame from most National Championships to least. (Line 160) 
Most Common from Simulation (Line 163)

```{r}
tbl <- sort(table(nat.champ.mat$National_Champ), decreasing = TRUE)
tbl

t(t(apply(nat.champ.mat, 2, function(x){names(sort(table(x), decreasing = TRUE))[1]})))
```
Now visualize for more fun and make some good picks. 

National Champions
```{r}
barplot(tbl/10000, col="cyan3", las=2, ylab="Probability", main="Win The Ship")
```

Final four probability by region
```{r}
south.tbl <- sort(table(nat.champ.mat$South.Regional_Final), decreasing = TRUE)
west.tbl <- sort(table(nat.champ.mat$West.Regional_Final), decreasing = TRUE)
east.tbl <- sort(table(nat.champ.mat$East.Regional_Final), decreasing = TRUE)
midwest.tbl <- sort(table(nat.champ.mat$Midwest.Regional_Final), decreasing = TRUE)

par(mfrow=c(2,2))
par(mar=c(8,4,4,4))

barplot(south.tbl/10000, col="blue", las=2, ylab="Probability", main="South Region")
barplot(west.tbl/10000, col="green", las=2, ylab="Probability", main="West Region")
barplot(midwest.tbl/10000, col="red", las=2, ylab="Probability", main="Midwest Region")
barplot(east.tbl/10000, col="purple", las=2, ylab="Probability", main="East Region")
```


2 vs 15
```{r}
iowast.sdak.tbl <- sort(table(nat.champ.mat$East.First_Round2), decreasing = TRUE)
zona.lbsu.tbl <- sort(table(nat.champ.mat$West.First_Round2 ), decreasing = TRUE)
marq.wky.tbl <- sort(table(nat.champ.mat$South.First_Round2 ), decreasing = TRUE)
tenn.spu.tbl <- sort(table(nat.champ.mat$Midwest.First_Round2 ), decreasing = TRUE)

barplot(iowast.sdak.tbl/10000, col=c("firebrick", "blue"), ylab="Probability")
barplot(zona.lbsu.tbl/10000, col=c("red","gold"), ylab="Probability")
barplot(marq.wky.tbl/10000, col=c("darkblue", "red"), ylab="Probability")
barplot(tenn.spu.tbl/10000, col=c("orange","deepskyblue4"), ylab="Probability")
```

3 vs 14
```{r}
ill.more.tbl <- sort(table(nat.champ.mat$East.First_Round3), decreasing = TRUE)
bay.colg.tbl <- sort(table(nat.champ.mat$West.First_Round3 ), decreasing = TRUE)
uk.oak.tbl <- sort(table(nat.champ.mat$South.First_Round3 ), decreasing = TRUE)
crei.akro.tbl <- sort(table(nat.champ.mat$Midwest.First_Round3 ), decreasing = TRUE)

barplot(ill.more.tbl/10000, col=c("chocolate1","royalblue1"), ylab="Probability")
barplot(bay.colg.tbl/10000, col=c("forestgreen","firebrick2"), ylab="Probability")
barplot(uk.oak.tbl/10000, col=c("blue","gold3"), ylab="Probability")
barplot(crei.akro.tbl/10000, col=c("blue1","darkblue"), ylab="Probability")
```

4 vs 13
```{r}
aub.yale.tbl <- sort(table(nat.champ.mat$East.First_Round4), decreasing = TRUE)
bama.char.tbl <- sort(table(nat.champ.mat$West.First_Round4 ), decreasing = TRUE)
duke.vrm.tbl <- sort(table(nat.champ.mat$South.First_Round4), decreasing = TRUE)
ku.sam.tbl <- sort(table(nat.champ.mat$Midwest.First_Round4), decreasing = TRUE)

barplot(aub.yale.tbl/10000, col=c("tomato2","royalblue4"), ylab="Probability")
barplot(bama.char.tbl/10000, col=c("red4","indianred4"), ylab="Probability")
barplot(duke.vrm.tbl/10000, col=c("blue","gold"), ylab="Probability")
barplot(ku.sam.tbl/10000, col=c("blue","red"), ylab="Probability")
```

5 vs 12
```{r}
sdsu.uab.tbl <- sort(table(nat.champ.mat$East.First_Round5 ), decreasing = TRUE)
mary.gcu.tbl <- sort(table(nat.champ.mat$West.First_Round5 ), decreasing = TRUE)
wis.jmu.tbl <- sort(table(nat.champ.mat$South.First_Round5 ), decreasing = TRUE)
zaga.mcns.tbl <- sort(table(nat.champ.mat$Midwest.First_Round5 ), decreasing = TRUE)

barplot(sdsu.uab.tbl/10000, col=c("darkred","darkgreen"), ylab="Probability")
barplot(mary.gcu.tbl/10000, col=c("red3","purple"), ylab="Probability")
barplot(wis.jmu.tbl/10000, col=c("red","darkblue"), ylab="Probability")
barplot(zaga.mcns.tbl/10000, col=c("grey30","gold"), ylab="Probability")
```

6 vs 11
```{r}
byu.dnq.tbl <- sort(table(nat.champ.mat$East.First_Round6 ), decreasing = TRUE)
clem.nmu.tbl <- sort(table(nat.champ.mat$West.First_Round6 ), decreasing = TRUE)
tt.ncst.tbl <- sort(table(nat.champ.mat$South.First_Round6 ), decreasing = TRUE)
ore.scu.tbl <- sort(table(nat.champ.mat$Midwest.First_Round6 ), decreasing = TRUE)

barplot(byu.dnq.tbl/10000, col=c("blue","darkblue"), ylab="Probability")
barplot(clem.nmu.tbl/10000, col=c("orange2","darkred"), ylab="Probability")
barplot(tt.ncst.tbl/10000, col=c("red","red3"), ylab="Probability")
barplot(ore.scu.tbl/10000, col=c("darkred","green"), ylab="Probability")
```

7 vs 10
```{r}
wazzu.drk.tbl <- sort(table(nat.champ.mat$East.First_Round7 ), decreasing = TRUE)
day.nev.tbl <- sort(table(nat.champ.mat$West.First_Round7 ), decreasing = TRUE)
fla.colo.bsu.tbl <- sort(table(nat.champ.mat$South.First_Round7 ), decreasing = TRUE)
tex.csu.uva.tbl <- sort(table(nat.champ.mat$Midwest.First_Round7 ), decreasing = TRUE)

barplot(wazzu.drk.tbl/10000, col=c("blue","grey"), ylab="Probability")
barplot(day.nev.tbl/10000, col=c("darkblue","red"), ylab="Probability")
barplot(fla.colo.bsu.tbl/10000, col=c("blue","gold","orange"), ylab="Probability")
barplot(tex.csu.uva.tbl/10000, col=c("orange3","darkblue", "darkgreen"), ylab="Probability")
```

8 vs 9
```{r}
fau.nw.tbl <- sort(table(nat.champ.mat$East.First_Round8), decreasing = TRUE)
misst.msu.tbl <- sort(table(nat.champ.mat$West.First_Round8 ), decreasing = TRUE)
neb.tam.tbl <- sort(table(nat.champ.mat$South.First_Round8 ), decreasing = TRUE)
usu.tcu.tbl <- sort(table(nat.champ.mat$Midwest.First_Round8 ), decreasing = TRUE)

barplot(fau.nw.tbl/10000, col=c("royalblue","purple3"), ylab="Probability")
barplot(misst.msu.tbl/10000, col=c("green4","red4"), ylab="Probability")
barplot(neb.tam.tbl/10000, col=c("darkred","red"), ylab="Probability")
barplot(usu.tcu.tbl/10000, col=c("purple","royalblue3"), ylab="Probability")
```


25 Bracket Simulations 
```{r}
sim1 <- March.Madness.Simulation(dat=dat24)
sim2 <- March.Madness.Simulation(dat=dat24)
sim3 <- March.Madness.Simulation(dat=dat24)
sim4 <- March.Madness.Simulation(dat=dat24)
sim5 <- March.Madness.Simulation(dat=dat24)
sim6 <- March.Madness.Simulation(dat=dat24)
sim7 <- March.Madness.Simulation(dat=dat24)
sim8 <- March.Madness.Simulation(dat=dat24)
sim9 <- March.Madness.Simulation(dat=dat24)
sim10 <- March.Madness.Simulation(dat=dat24)
sim11 <- March.Madness.Simulation(dat=dat24)
sim12 <- March.Madness.Simulation(dat=dat24)
sim13 <- March.Madness.Simulation(dat=dat24)
sim14 <- March.Madness.Simulation(dat=dat24)
sim15 <- March.Madness.Simulation(dat=dat24)
sim16 <- March.Madness.Simulation(dat=dat24)
sim17 <- March.Madness.Simulation(dat=dat24)
sim18 <- March.Madness.Simulation(dat=dat24)
sim19 <- March.Madness.Simulation(dat=dat24)
sim20 <- March.Madness.Simulation(dat=dat24)
sim21 <- March.Madness.Simulation(dat=dat24)
sim22 <- March.Madness.Simulation(dat=dat24)
sim23 <- March.Madness.Simulation(dat=dat24)
sim24 <- March.Madness.Simulation(dat=dat24)
sim25 <- March.Madness.Simulation(dat=dat24)
```
```{r}
sim26 <- March.Madness.Simulation(dat=dat24)
sim27 <- March.Madness.Simulation(dat=dat24)
sim28 <- March.Madness.Simulation(dat=dat24)
sim29 <- March.Madness.Simulation(dat=dat24)
sim30 <- March.Madness.Simulation(dat=dat24)
sim31 <- March.Madness.Simulation(dat=dat24)
sim32 <- March.Madness.Simulation(dat=dat24)
sim33 <- March.Madness.Simulation(dat=dat24)
sim34 <- March.Madness.Simulation(dat=dat24)
sim35 <- March.Madness.Simulation(dat=dat24)
sim36 <- March.Madness.Simulation(dat=dat24)
sim37 <- March.Madness.Simulation(dat=dat24)
sim38 <- March.Madness.Simulation(dat=dat24)
sim39 <- March.Madness.Simulation(dat=dat24)
sim40 <- March.Madness.Simulation(dat=dat24)
sim41 <- March.Madness.Simulation(dat=dat24)
sim42 <- March.Madness.Simulation(dat=dat24)
sim43 <- March.Madness.Simulation(dat=dat24)
sim44 <- March.Madness.Simulation(dat=dat24)
sim45 <- March.Madness.Simulation(dat=dat24)
sim46 <- March.Madness.Simulation(dat=dat24)
sim47 <- March.Madness.Simulation(dat=dat24)
sim48 <- March.Madness.Simulation(dat=dat24)
sim49 <- March.Madness.Simulation(dat=dat24)
sim50<- March.Madness.Simulation(dat=dat24)
sim51 <- March.Madness.Simulation(dat=dat24)
sim52 <- March.Madness.Simulation(dat=dat24)
sim53 <- March.Madness.Simulation(dat=dat24)
sim54 <- March.Madness.Simulation(dat=dat24)
sim55 <- March.Madness.Simulation(dat=dat24)
sim56 <- March.Madness.Simulation(dat=dat24)
sim57 <- March.Madness.Simulation(dat=dat24)
sim58 <- March.Madness.Simulation(dat=dat24)
sim59 <- March.Madness.Simulation(dat=dat24)
sim60 <- March.Madness.Simulation(dat=dat24)
sim61 <- March.Madness.Simulation(dat=dat24)
sim62 <- March.Madness.Simulation(dat=dat24)
sim63 <- March.Madness.Simulation(dat=dat24)
sim64 <- March.Madness.Simulation(dat=dat24)
sim65 <- March.Madness.Simulation(dat=dat24)
sim66 <- March.Madness.Simulation(dat=dat24)
sim67 <- March.Madness.Simulation(dat=dat24)
sim68 <- March.Madness.Simulation(dat=dat24)
sim69 <- March.Madness.Simulation(dat=dat24)
sim70 <- March.Madness.Simulation(dat=dat24)
sim71 <- March.Madness.Simulation(dat=dat24)
sim72 <- March.Madness.Simulation(dat=dat24)
sim73 <- March.Madness.Simulation(dat=dat24)
sim74 <- March.Madness.Simulation(dat=dat24)
sim75 <- March.Madness.Simulation(dat=dat24)
sim76 <- March.Madness.Simulation(dat=dat24)
sim77 <- March.Madness.Simulation(dat=dat24)
sim78 <- March.Madness.Simulation(dat=dat24)
sim79 <- March.Madness.Simulation(dat=dat24)
sim80 <- March.Madness.Simulation(dat=dat24)
sim81 <- March.Madness.Simulation(dat=dat24)
sim82 <- March.Madness.Simulation(dat=dat24)
sim83 <- March.Madness.Simulation(dat=dat24)
sim84 <- March.Madness.Simulation(dat=dat24)
sim85 <- March.Madness.Simulation(dat=dat24)
sim86 <- March.Madness.Simulation(dat=dat24)
```

```{r}
sim87 <- March.Madness.Simulation(dat=dat24)
sim88 <- March.Madness.Simulation(dat=dat24)
sim89 <- March.Madness.Simulation(dat=dat24)
sim90 <- March.Madness.Simulation(dat=dat24)
sim91 <- March.Madness.Simulation(dat=dat24)
sim92 <- March.Madness.Simulation(dat=dat24)
sim93 <- March.Madness.Simulation(dat=dat24)
sim94 <- March.Madness.Simulation(dat=dat24)
sim95 <- March.Madness.Simulation(dat=dat24)
sim96 <- March.Madness.Simulation(dat=dat24)
sim97 <- March.Madness.Simulation(dat=dat24)
sim98 <- March.Madness.Simulation(dat=dat24)
sim99 <- March.Madness.Simulation(dat=dat24)
```


